### 2015-11-30 paper/paper.Rmd
start

---
title: R Packages to Aid in Handling Web Access Logs
author:
  - name: Oliver Keyes
    affiliation: Wikimedia Foundation
    address:
    - line 1
    - line 2
    email:  author1@work
  - name: Bob Rudis
    affiliation: Rapid7
    address:
    - line 1
    - line 2
    email:  author2@work
abstract: >
  Web access logs contain information on HTTP(S) requests and form a key part of both industry and academic explorations
  of human behaviour on the internet - explorations commonly performed in R. In this paper we explain and demonstrate
  a series of packages - **webreadr**, **urltools**, **iptools** and **rgeolocate** - designed to efficiently read in, parse
  and munge access log data, allowing researchers to handle it easily.
preamble: >
output: rticles::rjournal_article
---

## Introduction

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

## Reading access logs

This section may contain a figure such as Figure \ref{figure:rlogo}.

\begin{figure}[htbp]
  \centering
  \includegraphics{Rlogo}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

## Decoding and parsing URLs

There will likely be several sections, perhaps including code snippets, such as:

```{r}
x <- 1:10
x
```

## Summary

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](http://journal.r-project.org/latex/RJauthorguide.pdf).

\bibliography{RJreferences}

### 2015-11-30 paper/paper.tex
start

% !TeX root = RJwrapper.tex
\title{R Packages to Aid in Handling Web Access Logs}
\author{by Oliver Keyes, Bob Rudis}

\maketitle

\abstract{
Web access logs contain information on HTTP(S) requests and form a key
part of both industry and academic explorations of human behaviour on
the internet - explorations commonly performed in R. In this paper we
explain and demonstrate a series of packages - \emph{webreadr},
\emph{urltools}, \emph{iptools} and \emph{rgeolocate} - designed to
efficiently read in, parse and munge access log data, allowing
researchers to handle it easily.
}

\subsection{Introduction}\label{introduction}

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

\subsection{Section title in sentence
case}\label{section-title-in-sentence-case}

This section may contain a figure such as Figure \ref{figure:rlogo}.

\begin{figure}[htbp]
  \centering
  \includegraphics{Rlogo}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

\subsection{Another section}\label{another-section}

There will likely be several sections, perhaps including code snippets,
such as:

\begin{Schunk}
\begin{Sinput}
x <- 1:10
x
\end{Sinput}
\begin{Soutput}
#>  [1]  1  2  3  4  5  6  7  8  9 10
\end{Soutput}
\end{Schunk}

\subsection{Summary}\label{summary}

This file is only a basic article template. For full details of
\emph{The R Journal} style and information on how to prepare your
article for submission, see the
\href{http://journal.r-project.org/latex/RJauthorguide.pdf}{Instructions
for Authors}.

\bibliography{RJreferences}

\address{
Oliver Keyes\\
Wikimedia Foundation\\
line 1\\ line 2\\
}
\href{mailto:author1@work}{\nolinkurl{author1@work}}

\address{
Bob Rudis\\
Rapid7\\
line 1\\ line 2\\
}
\href{mailto:author2@work}{\nolinkurl{author2@work}}

### 2015-11-30 urltools_benchmarks.R

@@ -1,0 +1,36 @@
+# This code generates visualisations and benchmarks highlighting the difference in speed between urltools and non-urltools
+# equivalent functions, for inclusion in the paper "R Packages to Aid in Handling Web Access Logs"
+
+library(httr) # only existing parser equivalent
+library(microbenchmark) # benchmarking
+library(ggplot2) # Plotting benchmarks
+library(wmf) # theming plots
+library(urltools) # doy
+
+# First, decoding. This is fairly simple.
+# Set up the functions
+urltools_decode <- urltools::url_decode
+r_base_decode   <- function(x){
+  vapply(x, URLdecode, "character")
+}
+
+# Generate the data. We want a moderately complex but not unnecessarily so URL, and we want 1m of them.
+# Let's use the urltools example
+example_urls <- rep("https://en.wikipedia.org/wiki/File:Vice_City_Public_Radio_%28logo%29.jpg", 1000000)
+
+# Generate benchmarks
+decoding_benchmarks <- microbenchmark({(urltools_decode(example_urls))},
+                                      {(r_base_decode(example_urls))})
+
+# Test parsing. Also fairly simple
+# Set up the functions:
+urltools_parse <- urltools::url_parse
+httr_parse     <- function(x){
+  lapply(x, httr::parse_url)
+}
+
+parsing_benchmarks <- microbenchmark({(urltools_parse(example_urls))},
+                                     {(httr_parse(example_urls))})
+
+# Save the results to disc, because nobody wants to lose benchmarks that take _this long_ to compute.
+save(decoding_benchmarks, parsing_benchmarks, file = "urltools_benchmarks.RData")

### 2015-11-30 webreadr_benchmarks.R

@@ -1,0 +1,27 @@
+# This code generates visualisations and benchmarks highlighting the difference in speed between webreadr and base R
+# equivalent functions, for inclusion in the paper "R Packages to Aid in Handling Web Access Logs"
+
+library(httr) # only existing parser equivalent
+library(microbenchmark) # benchmarking
+library(ggplot2) # Plotting benchmarks
+library(wmf) # theming plots
+library(webreadr) # doy
+
+# We already have an example file, expanded_squid_log.squid, containing 600000 values. Let's use that as a basis.
+
+# Create the functions
+webreadr_read_squid <- webreadr::read_squid
+r_base_read_squid <- function(filename){
+  names <- c("timestamp", "time_elapsed", "ip_address", "status_code",
+             "bytes_sent","http_method", "url","remote_user_ident","peer_info")
+  data <- read.delim(filename, sep = " ", header = FALSE, as.is = TRUE, col.names = names)
+  data$timestamp <- as.POSIXlt(data$timestamp, origin = "1970-01-01", tz = "UTC")
+  return(data)
+}
+
+# Benchmark
+reading_benchmarks <- microbenchmark({(webreadr_read_squid("expanded_squid_log.squid"))},
+                                     {(r_base_read_squid("expanded_squid_log.squid"))})
+
+# Save benchmarks
+save(reading_benchmarks, file = "webreadr_benchmarks.RData")

### 2015-12-03 paper/paper.Rmd
add benchmarks

@@
 
-Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
+The rise of the World Wide Web over the last few decades has made it dramatically easier to access and transfer data, and R \citep{RCore} boasts abundant functionality when it comes to taking data *from* the web.
+Base R itself has simple file downloading and page reading capabilities, through the `download.file` and `readLines` functions, and additional functionality is made available for handling web-accessible data through packages such as **httr** \citep{httr}.
+
+Data *on* the web is not, however, the only kind of web data that interests researchers; web traffic is, in and of itself, an interesting data source.
+Access logs - records of connections between users and a web server - are an asset and resource for people studying everything from user behaviour on the internet \citep{halfak}, to website performance \citep{performance}, to information security \citep{infosec}.
+
+As a statistically-oriented programming language, R is commonly used by these same researchers for data analysis, testing and reporting - but it lacks tools designed for the kinds of data sources and formats that are encountered when dealing with access logs.
+In this article we review the use cases for particular operations over web data, the limitations in base R when it comes to performing those operations, and a suite of R packages designed to overcome them: **webreadr** \citep{webreadr}, designed for reading access logs in,
+**urltools** \citep{urltools}, for URL manipulation, **iptools** \citep{rgeolocate} for handling IP addresses, and **rgeolocate** \citep{rgeolocate} for direct IP geolocation.
 
@@
 
-This section may contain a figure such as Figure \ref{figure:rlogo}.
+The first task with any data analysis is to read the data into R so that it can be manipulated.
+With access logs this is slightly complicated by the fact that there is no one standard for what a log should look like; instead, there are multiple competing approaches from different software platforms and eras.
+These include the Common Log Format (CLF), the confusingly-named Combined Log Format, and formats used by individual, commonly-used software platforms - such as the custom format for the Squid internet caching software, and the format used by Amazon Web Services (AWS).
+
+The difference between formats can easily be shown by looking at how timestamps are represented:
+
+\begin{table}[ht]
+\centering
+\caption{Timestamps in Common Access Log Formats}
+\begin{tabular}{rlr}
+  \hline
+Log Type & Timestamp Columns & Timestamp Format \\ 
+  \hline
+Common Log Format & 1 & 10/Oct/2000:13:55:36 -0700\\ 
+Combined Log Format & 1 & 26/Apr/2000:00:23:48 -0400\\ 
+Squid & 1 & 1286536309.450\\ 
+AWS & 2 & 2014-05-23	01:13:11\\ 
+   \hline
+\end{tabular}
+\end{table}
+
+With four log types, we have three different timestamp formats - and that's only one of the many columns that could appear.
+These logs also vary in whether they specify quoting fields (or sanitising unquoted ones), the columns they contain and the data each column contains in turn.
+
+To make reading access logs into R as easy as possible we created the **webreadr** package.
+This contains user-friendly equivalents to `read.table` for each type of log, detecting the fields that should appear, converting the timestamps into POSIX objects, and merging fields or splitting fields where necessary.
+The package contains four core functions, one for each form of log: `read_clf` for the Common Log Format, `read_combined` for the Combined Log Format, and `read_squid` and `read_aws` for Squid and AWS formats respectively.
+Each one abstracts away the complexity of specifying column names and formats, and instead allows a researcher to read a file in with a minimal amount of work.
+
+As the name suggests, it is built not on top of base R but on top of the **readr** \citep{readr} package, allowing us to take advantage of substantial speed improvements that package's base functions have over base R.
+These improvements can be seen in the visualisation below, which uses **microbenchmark** \citep{microbenchmark} to compare 100 reads of a 600,000-line "squid" formatted file with webreadr to the same operation performed in base R:
 
@@
 
+**webreadr**
+
 ## Decoding and parsing URLs
 
-There will likely be several sections, perhaps including code snippets, such as:
+Within access logs, URLs appear - both to describe the web asset or page that the user requested and where the user came from.
+These fields are usually named <code>url</code> and <code>referer</code> respectively.
 
-```{r}
-x <- 1:10
-x
-```
+### Decoding
+Both values can be percent-encoded, allowing them toinclude characters that are valid but reserved by the URL specification as having special meanings ("reserved characters").
+A `#` symbol, for example, is encoded as <code>%23</code>: a percentage symbol, followed by the byte-code for that character.
+
+The encoding of reserved characters is useful, since it means that URL paths and queries can contain a vast array of values - but it makes data analysis tougher to do.
+Examples of common data analysis or cleaning operations that become more difficult are:
+
+1. **Aggregation**. Aggregating URLs together is useful to identify, for example, the relative usage and popularity of particular pages
+in your data - but it becomes tougher if encoding is inconsistent, because two URLs could hold the same value but *look* very different.
+2.
+**Value selection**.
+With text-based data, regular expressions are a common way of filtering or selecting entries that meet particular conditions, but things become fuzzy when you have to look not just for particular characters (a space, say) but also the encoded values (<code>%20</code>).
+3.
+**Exploratory data analysis** (EDA).
+EDA is a common initial step to investigate a data set, examining the variables and values it contains and whether they meet a researcher's expectations - but on a practical basis it becomes difficult when the values aren't human-readable.
+
+The solution is to be able to consistently decode URLs, which makes URL-based data far easier to analyse.
+Base R does contain a function, <code>URLdecode</code>, for doing just that, but is neither vectorised nor based on compiled code; with large datasets it takes an extremely long time.
+
+To solve this common problem in analysing request logs, the **urltools** [@urltools] package was created. This contains a function,
+<code>url_decode</code>, which decodes URLs - but relies on vectorised, compiled code.
+Benchmarking the two approaches against each other shows that the `urltools` implementation is approximately 60-70 times faster over large datasets.
+Again using **microbenchmark**, if we compare the vectorised decoding of 1,000,000 URLs with `urltools` against `URLdecode` in a `vapply` loop, we see a 60-70 times speed improvement of urltools over base R:
 

### 2015-12-03 paper/paper.tex
add benchmarks

@@
 the internet - explorations commonly performed in R. In this paper we
-explain and demonstrate a series of packages - \emph{webreadr},
-\emph{urltools}, \emph{iptools} and \emph{rgeolocate} - designed to
-efficiently read in, parse and munge access log data, allowing
+explain and demonstrate a series of packages - \textbf{webreadr},
+\textbf{urltools}, \textbf{iptools} and \textbf{rgeolocate} - designed
+to efficiently read in, parse and munge access log data, allowing
 researchers to handle it easily.
@@
 
-Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
+The rise of the World Wide Web over the last few decades has made it
+dramatically easier to access and transfer data, and R \citep{RCore}
+boasts abundant functionality when it comes to taking data \emph{from}
+the web. Base R itself has simple file downloading and page reading
+capabilities, through the \texttt{download.file} and \texttt{readLines}
+functions, and additional functionality is made available for handling
+web-accessible data through packages such as \textbf{httr} \citep{httr}.
 
-\subsection{Section title in sentence
-case}\label{section-title-in-sentence-case}
+Data \emph{on} the web is not, however, the only kind of web data that
+interests researchers; web traffic is, in and of itself, an interesting
+data source. Access logs - records of connections between users and a
+web server - are an asset and resource for people studying everything
+from user behaviour on the internet \citep{halfak}, to website
+performance \citep{performance}, to information security
+\citep{infosec}.
 
-This section may contain a figure such as Figure \ref{figure:rlogo}.
+As a statistically-oriented programming language, R is commonly used by
+these same researchers for data analysis, testing and reporting - but it
+lacks tools designed for the kinds of data sources and formats that are
+encountered when dealing with access logs. In this article we review the
+use cases for particular operations over web data, the limitations in
+base R when it comes to performing those operations, and a suite of R
+packages designed to overcome them: \textbf{webreadr} \citep{webreadr},
+designed for reading access logs in, \textbf{urltools} \citep{urltools},
+for URL manipulation, \textbf{iptools} \citep{rgeolocate} for handling
+IP addresses, and \textbf{rgeolocate} \citep{rgeolocate} for direct IP
+geolocation.
+
+\subsection{Reading access logs}\label{reading-access-logs}
+
+The first task with any data analysis is to read the data into R so that
+it can be manipulated. With access logs this is slightly complicated by
+the fact that there is no one standard for what a log should look like;
+instead, there are multiple competing approaches from different software
+platforms and eras. These include the Common Log Format (CLF), the
+confusingly-named Combined Log Format, and formats used by individual,
+commonly-used software platforms - such as the custom format for the
+Squid internet caching software, and the format used by Amazon Web
+Services (AWS).
+
+The difference between formats can easily be shown by looking at how
+timestamps are represented:
+
+\begin{table}[ht]
+\centering
+\caption{Timestamps in Common Access Log Formats}
+\begin{tabular}{rlr}
+  \hline
+Log Type & Timestamp Columns & Timestamp Format \\ 
+  \hline
+Common Log Format & 1 & 10/Oct/2000:13:55:36 -0700\\ 
+Combined Log Format & 1 & 26/Apr/2000:00:23:48 -0400\\ 
+Squid & 1 & 1286536309.450\\ 
+AWS & 2 & 2014-05-23    01:13:11\\ 
+   \hline
+\end{tabular}
+\end{table}
+
+With four log types, we have three different timestamp formats - and
+that's only one of the many columns that could appear. These logs also
+vary in whether they specify quoting fields (or sanitising unquoted
+ones), the columns they contain and the data each column contains in
+turn.
+
+To make reading access logs into R as easy as possible we created the
+\textbf{webreadr} package. This contains user-friendly equivalents to
+\texttt{read.table} for each type of log, detecting the fields that
+should appear, converting the timestamps into POSIX objects, and merging
+fields or splitting fields where necessary. The package contains four
+core functions, one for each form of log: \texttt{read\_clf} for the
+Common Log Format, \texttt{read\_combined} for the Combined Log Format,
+and \texttt{read\_squid} and \texttt{read\_aws} for Squid and AWS
+formats respectively. Each one abstracts away the complexity of
+specifying column names and formats, and instead allows a researcher to
+read a file in with a minimal amount of work.
+
+As the name suggests, it is built not on top of base R but on top of the
+\textbf{readr} \citep{readr} package, allowing us to take advantage of
+substantial speed improvements that package's base functions have over
+base R. These improvements can be seen in the visualisation below, which
+uses \textbf{microbenchmark} \citep{microbenchmark} to compare 100 reads
+of a 600,000-line ``squid'' formatted file with webreadr to the same
+operation performed in base R:
 
@@
 
-\subsection{Another section}\label{another-section}
+\textbf{webreadr}
 
-There will likely be several sections, perhaps including code snippets,
-such as:
+\subsection{Decoding and parsing URLs}\label{decoding-and-parsing-urls}
 
-\begin{Schunk}
-\begin{Sinput}
-x <- 1:10
-x
-\end{Sinput}
-\begin{Soutput}
-#>  [1]  1  2  3  4  5  6  7  8  9 10
-\end{Soutput}
-\end{Schunk}
+Within access logs, URLs appear - both to describe the web asset or page
+that the user requested and where the user came from. These fields are
+usually named url and referer respectively.
+
+\subsubsection{Decoding}\label{decoding}
+
+Both values can be percent-encoded, allowing them toinclude characters
+that are valid but reserved by the URL specification as having special
+meanings (``reserved characters''). A \texttt{\#} symbol, for example,
+is encoded as \%23: a percentage symbol, followed by the byte-code for
+that character.
+
+The encoding of reserved characters is useful, since it means that URL
+paths and queries can contain a vast array of values - but it makes data
+analysis tougher to do. Examples of common data analysis or cleaning
+operations that become more difficult are:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Aggregation}. Aggregating URLs together is useful to identify,
+  for example, the relative usage and popularity of particular pages in
+  your data - but it becomes tougher if encoding is inconsistent,
+  because two URLs could hold the same value but \emph{look} very
+  different.
+\item
+  \textbf{Value selection}. With text-based data, regular expressions
+  are a common way of filtering or selecting entries that meet
+  particular conditions, but things become fuzzy when you have to look
+  not just for particular characters (a space, say) but also the encoded
+  values (\%20).
+\item
+  \textbf{Exploratory data analysis} (EDA). EDA is a common initial step
+  to investigate a data set, examining the variables and values it
+  contains and whether they meet a researcher's expectations - but on a
+  practical basis it becomes difficult when the values aren't
+  human-readable.
+\end{enumerate}
+
+The solution is to be able to consistently decode URLs, which makes
+URL-based data far easier to analyse. Base R does contain a function,
+URLdecode, for doing just that, but is neither vectorised nor based on
+compiled code; with large datasets it takes an incredibly long time.
 

### 2015-12-03 urltools_benchmarks.R

@@ -22,6 +22,15 @@ example_urls <- rep("https://en.wikipedia.org/wiki/File:Vice_City_Public_Radio_%
 decoding_benchmarks <- microbenchmark({(urltools_decode(example_urls))},
                                       {(r_base_decode(example_urls))})
 
+# Reset levels
+levels(decoding_benchmarks$expr)[grepl(x = levels(decoding_benchmarks$expr), pattern = "urltools")] <- "urltools"
+levels(decoding_benchmarks$expr)[levels(decoding_benchmarks$expr) != "urltools"] <- "base R"
+
+# Plot
+ggsave(file = "decoding_benchmarks.png",
+       plot = autoplot(decoding_benchmarks) + theme_fivethirtynine(base_size = 14) +
+         labs(title = "Decoding 1m URLs, base R versus urltools (log-scaled)"))
+
 # Test parsing. Also fairly simple
 # Set up the functions:
 urltools_parse <- urltools::url_parse
@@ -29,8 +38,19 @@ httr_parse     <- function(x){
   lapply(x, httr::parse_url)
 }
 
+# Run
 parsing_benchmarks <- microbenchmark({(urltools_parse(example_urls))},
                                      {(httr_parse(example_urls))})
 
+# Reset levels
+levels(parsing_benchmarks$expr)[grepl(x = levels(parsing_benchmarks$expr), pattern = "urltools")] <- "urltools"
+levels(parsing_benchmarks$expr)[levels(parsing_benchmarks$expr) != "urltools"] <- "httr"
+
+# Plot
+ggsave(file = "parsing_benchmarks.png",
+       plot = autoplot(parsing_benchmarks) + theme_fivethirtynine(base_size = 14) +
+         labs(title = "Parsing 1m URLs, httr versus urltools (log-scaled)"))
+
 # Save the results to disc, because nobody wants to lose benchmarks that take _this long_ to compute.
 save(decoding_benchmarks, parsing_benchmarks, file = "urltools_benchmarks.RData")
+

### 2015-12-03 webreadr_benchmarks.R

@@ -23,5 +23,16 @@ r_base_read_squid <- function(filename){
 reading_benchmarks <- microbenchmark({(webreadr_read_squid("expanded_squid_log.squid"))},
                                      {(r_base_read_squid("expanded_squid_log.squid"))})
 
+# Relevel
+levels(reading_benchmarks$expr)[grepl(x = levels(reading_benchmarks$expr), pattern = "webreadr")] <- "webreadr"
+levels(reading_benchmarks$expr)[levels(reading_benchmarks$expr) != "webreadr"] <- "base R"
+
+# Plot
+ggsave(file = "reading_benchmarks.png",
+       autoplot(reading_benchmarks) + theme_fivethirtynine(base_size = 14)  + theme_fivethirtynine(base_size = 14) + 
+         scale_y_continuous(expand=c(0,0)) + 
+         labs(y = "Time [seconds]", title = "Reading 600k lines of access log, base R versus webreadr"))
+
 # Save benchmarks
 save(reading_benchmarks, file = "webreadr_benchmarks.RData")
+

### 2015-12-03 paper/paper.Rmd
update

@@
 ---
-
 ## Introduction
@@
 
-\begin{figure}[htbp]
+\begin{figure}[h]
     \centering
-  \includegraphics{Rlogo}
-  \caption{The logo of R.}
-  \label{figure:rlogo}
+    \includegraphics[scale=0.4]{reading_benchmarks}
+    \caption{Results of microbenchmark run: read\char`_squid versus base-R equivalent code}
 \end{figure}
 
-**webreadr**
+**webreadr** takes 1.2 seconds to read this file in at a minimum, and 1.48 at maximum, with a median of 1.33.
+This is consistently 3.5-6 times faster than the equivalent base R functionality - and is also, as explained, far easier to use.
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.4]{decoding_benchmarks}
+    \caption{Results of microbenchmark run: url\char`_decode versus base-R equivalent code}
+\end{figure}
+
 ## Summary

### 2015-12-03 paper/paper.tex
update

@@
 
-\begin{figure}[htbp]
+\begin{figure}[h]
     \centering
-  \includegraphics{Rlogo}
-  \caption{The logo of R.}
-  \label{figure:rlogo}
+    \includegraphics[scale=0.4]{reading_benchmarks}
+    \caption{Results of microbenchmark run: read\char`_squid versus base-R equivalent code}
 \end{figure}
 
-\textbf{webreadr}
+\textbf{webreadr} takes 1.2 seconds to read this file in at a minimum,
+and 1.48 at maximum, with a median of 1.33. This is consistently 3.5-6
+times faster than the equivalent base R functionality - and is also, as
+explained, far easier to use.
 
@@
 URLdecode, for doing just that, but is neither vectorised nor based on
-compiled code; with large datasets it takes an incredibly long time.
+compiled code; with large datasets it takes an extremely long time.
+
+To solve this common problem in analysing request logs, the
+\textbf{urltools} {[}@urltools{]} package was created. This contains a
+function, url\_decode, which decodes URLs - but relies on vectorised,
+compiled code. Benchmarking the two approaches against each other shows
+that the \texttt{urltools} implementation is approximately 60-70 times
+faster over large datasets. Again using \textbf{microbenchmark}, if we
+compare the vectorised decoding of 1,000,000 URLs with \texttt{urltools}
+against \texttt{URLdecode} in a \texttt{vapply} loop, we see a 60-70
+times speed improvement of urltools over base R:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.4]{decoding_benchmarks}
+    \caption{Results of microbenchmark run: url\char`_decode versus base-R equivalent code}
+\end{figure}
 

### 2015-12-03 urltools_benchmarks.R

@@ -27,7 +27,7 @@ levels(decoding_benchmarks$expr)[grepl(x = levels(decoding_benchmarks$expr), pat
 levels(decoding_benchmarks$expr)[levels(decoding_benchmarks$expr) != "urltools"] <- "base R"
 
 # Plot
-ggsave(file = "decoding_benchmarks.png",
+ggsave(file = "./paper/decoding_benchmarks.png",
        plot = autoplot(decoding_benchmarks) + theme_fivethirtynine(base_size = 14) +
          labs(title = "Decoding 1m URLs, base R versus urltools (log-scaled)"))
 
@@ -47,7 +47,7 @@ levels(parsing_benchmarks$expr)[grepl(x = levels(parsing_benchmarks$expr), patte
 levels(parsing_benchmarks$expr)[levels(parsing_benchmarks$expr) != "urltools"] <- "httr"
 
 # Plot
-ggsave(file = "parsing_benchmarks.png",
+ggsave(file = "./paper/parsing_benchmarks.png",
        plot = autoplot(parsing_benchmarks) + theme_fivethirtynine(base_size = 14) +
          labs(title = "Parsing 1m URLs, httr versus urltools (log-scaled)"))
 

### 2015-12-03 webreadr_benchmarks.R

@@ -28,7 +28,7 @@ levels(reading_benchmarks$expr)[grepl(x = levels(reading_benchmarks$expr), patte
 levels(reading_benchmarks$expr)[levels(reading_benchmarks$expr) != "webreadr"] <- "base R"
 
 # Plot
-ggsave(file = "reading_benchmarks.png",
+ggsave(file = "./paper/reading_benchmarks.png",
        autoplot(reading_benchmarks) + theme_fivethirtynine(base_size = 14)  + theme_fivethirtynine(base_size = 14) + 
          scale_y_continuous(expand=c(0,0)) + 
          labs(y = "Time [seconds]", title = "Reading 600k lines of access log, base R versus webreadr"))

### 2015-12-05 paper/paper.Rmd
add

@@
 
+### Parsing
+
+The standard for URLs \citep{RFC1738} divides them into a heirarchical sequence of components - the scheme ("http"), host ("en.wikipedia.org"), port ("800"), path ("wiki/Main_Page") and searchpart, or parameters ("action=edit").
+Together, these make
+up a URL ("http://en.wikipedia.org:800/wiki/Main_Page?action=edit").
+
+Parsing URLs to isolate and extract these components is a useful ability when it comes to exploring request logs; it lets a researcher pick out particular schemes, paths, hosts or other components to aggregate by, identifying how users are behaving and what they are visiting.
+It makes anonymising data - by removing, for example, the parameters or path, which can contain relatively unique information - easier.
+
+Base R does not have native code to parse URLs, but the **httr** package \citep{httr} contains a function, `parse_url`, designed to do just that and built on base R's capabilities.
+This function is neither vectorised nor based on compiled code, and produces a list, making looping over a set of URLs to parse each one a somewhat thorny experience.
+This is totally understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse request logs *en masse*.
+
+`urltools` contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame.
+Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, not having designed to be, does not.
+Indeed, benchmarking showed that httr's version is (for large vectors, at least) approximately 570 times slower than `url_parse`:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.4]{decoding_benchmarks}
+    \caption{Results of microbenchmark run: url\char`_parse versus httr equivalent code}
+\end{figure}
+
 ## Summary

### 2015-12-05 paper/paper.tex
add

@@
 
+\subsubsection{Parsing}\label{parsing}
+
+The standard for URLs \citep{RFC1738} divides them into a heirarchical
+sequence of components - the scheme (``http''), host
+(``en.wikipedia.org''), port (``800''), path (``wiki/Main\_Page'') and
+searchpart, or parameters (``action=edit''). Together, these make up a
+URL (``\url{http://en.wikipedia.org:800/wiki/Main_Page?action=edit}'').
+
+Parsing URLs to isolate and extract these components is a useful ability
+when it comes to exploring request logs; it lets a researcher pick out
+particular schemes, paths, hosts or other components to aggregate by,
+identifying how users are behaving and what they are visiting. It makes
+anonymising data - by removing, for example, the parameters or path,
+which can contain relatively unique information - easier.
+
+Base R does not have native code to parse URLs, but the \textbf{httr}
+package \citep{httr} contains a function, \texttt{parse\_url}, designed
+to do just that and built on base R's capabilities. This function is
+neither vectorised nor based on compiled code, and produces a list,
+making looping over a set of URLs to parse each one a somewhat thorny
+experience. This is totally understandable given the intent behind that
+function, which is to decompose individual URLs within the context of
+making HTTP requests, rather than to analyse request logs \emph{en
+masse}.
+
+\texttt{urltools} contains \texttt{url\_parse} - which does the same
+thing as the equivalent \texttt{httr} functionality, but in a vectorised
+way, relying on compiled code, and producing a data.frame. Within the
+context of parsing and processing access logs, this is far more useful,
+because it works efficiently over large sets: httr's functionality, not
+having designed to be, does not. Indeed, benchmarking showed that httr's
+version is (for large vectors, at least) approximately 570 times slower
+than \texttt{url\_parse}:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.4]{decoding_benchmarks}
+    \caption{Results of microbenchmark run: url\char`_parse versus httr equivalent code}
+\end{figure}
+
 \subsection{Summary}\label{summary}

### 2015-12-20 paper/paper.Rmd
a bit more

@@
 
-Base R does not have native code to parse URLs, but the **httr** package \citep{httr} contains a function, `parse_url`, designed to do just that and built on base R's capabilities.
-This function is neither vectorised nor based on compiled code, and produces a list, making looping over a set of URLs to parse each one a somewhat thorny experience.
-This is totally understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse request logs *en masse*.
+Base R does not have native code to parse URLs, but the **httr** package \citep{httr} contains a function, `parse_url`, designed to do just thatm built on R's regular expressions.
+This function is not vectorised, does not make use of compiled code internally, and produces a list rather than data.frame, making looping over a set of URLs to parse each one a time-consuming and awkward experience.
+This is understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse URLs *en masse*.
 
@@
 Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, not having designed to be, does not.
-Indeed, benchmarking showed that httr's version is (for large vectors, at least) approximately 570 times slower than `url_parse`:
+Indeed, benchmarking showed that `url_parse` is approximately 570 times faster than httr's equivalent function:
 
@@
     \centering
-    \includegraphics[scale=0.4]{decoding_benchmarks}
-    \caption{Results of microbenchmark run: url\char`_parse versus httr equivalent code}
+    \includegraphics[scale=0.4]{parsing_benchmarks}
+    \caption{Results of microbenchmark run: url\char`_parse versus httr's equivalent code}
 \end{figure}
+
+\newpage
+
+A vector of URLs passed into `url_parse` produces a data.frame, with one column for each of the IETF-supported components, and empty strings representing components that could not be found.
+Additionally, influenced by the style of the *lubridate* package \citep{lubridate}, `urltools` contains functions to get or *set* individual components:
+
+```{r, eval=TRUE}
+# Load urltools and construct a URL
+library(urltools)
+url <- "http://www.google.com/"
+
+# Get the scheme
+scheme(url)
+```
+
+```{r, eval=TRUE}
+# Set the scheme and observe the modification to the resulting URL
+scheme(url) <- "https"
+url
+```
+## IP manipulation and geolocation
 

### 2015-12-20 paper/paper.tex
a bit more

@@
 package \citep{httr} contains a function, \texttt{parse\_url}, designed
-to do just that and built on base R's capabilities. This function is
-neither vectorised nor based on compiled code, and produces a list,
-making looping over a set of URLs to parse each one a somewhat thorny
-experience. This is totally understandable given the intent behind that
-function, which is to decompose individual URLs within the context of
-making HTTP requests, rather than to analyse request logs \emph{en
-masse}.
+to do just thatm built on R's regular expressions. This function is not
+vectorised, does not make use of compiled code internally, and produces
+a list rather than data.frame, making looping over a set of URLs to
+parse each one a time-consuming and awkward experience. This is
+understandable given the intent behind that function, which is to
+decompose individual URLs within the context of making HTTP requests,
+rather than to analyse URLs \emph{en masse}.
 
@@
 because it works efficiently over large sets: httr's functionality, not
-having designed to be, does not. Indeed, benchmarking showed that httr's
-version is (for large vectors, at least) approximately 570 times slower
-than \texttt{url\_parse}:
+having designed to be, does not. Indeed, benchmarking showed that
+\texttt{url\_parse} is approximately 570 times faster than httr's
+equivalent function:
 
@@
     \centering
-    \includegraphics[scale=0.4]{decoding_benchmarks}
-    \caption{Results of microbenchmark run: url\char`_parse versus httr equivalent code}
+    \includegraphics[scale=0.4]{parsing_benchmarks}
+    \caption{Results of microbenchmark run: url\char`_parse versus httr's equivalent code}
 \end{figure}
+
+\newpage
+
+A vector of URLs passed into \texttt{url\_parse} produces a data.frame,
+with one column for each of the IETF-supported components, and empty
+strings representing components that could not be found. Additionally,
+influenced by the style of the \emph{lubridate} package
+\citep{lubridate}, \texttt{urltools} contains functions to get or set
+\emph{individual} components:
+
+\begin{Schunk}
+\begin{Sinput}
+library(urltools)
+url <- "http://www.google.com/"
+scheme(url)
+\end{Sinput}
+\begin{Soutput}
+#> [1] "http"
+\end{Soutput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+scheme(url) <- "https"
+url
+\end{Sinput}
+\begin{Soutput}
+#> [1] "https://www.google.com/"
+\end{Soutput}
+\end{Schunk}
+
+\subsection{IP manipulation and
+geolocation}\label{ip-manipulation-and-geolocation}
 

### 2015-12-20 paper/paper.Rmd
FIRST DRAFT DONE

@@
     email:  author2@work
+  - name: Jay Jacobs
+    affiliation: Magical Python Land
+    address:
+    - line 1
+    - line 2
+    email:  author3@work
 abstract: >
@@
   of human behaviour on the internet - explorations commonly performed in R. In this paper we explain and demonstrate
-  a series of packages - **webreadr**, **urltools**, **iptools** and **rgeolocate** - designed to efficiently read in, parse
+  a series of packages - \CRANpkg{webreadr}, \CRANpkg{urltools}, \CRANpkg{iptools} and \CRANpkg{rgeolocate} - designed to efficiently read in, parse
   and munge access log data, allowing researchers to handle it easily.
@@
 The rise of the World Wide Web over the last few decades has made it dramatically easier to access and transfer data, and R \citep{RCore} boasts abundant functionality when it comes to taking data *from* the web.
-Base R itself has simple file downloading and page reading capabilities, through the `download.file` and `readLines` functions, and additional functionality is made available for handling web-accessible data through packages such as **httr** \citep{httr}.
+Base R itself has simple file downloading and page reading capabilities, through the `download.file` and `readLines` functions, and additional functionality is made available for handling web-accessible data through packages such as \CRANpkg{httr} \citep{httr}.
 
@@
 As a statistically-oriented programming language, R is commonly used by these same researchers for data analysis, testing and reporting - but it lacks tools designed for the kinds of data sources and formats that are encountered when dealing with access logs.
-In this article we review the use cases for particular operations over web data, the limitations in base R when it comes to performing those operations, and a suite of R packages designed to overcome them: **webreadr** \citep{webreadr}, designed for reading access logs in,
-**urltools** \citep{urltools}, for URL manipulation, **iptools** \citep{rgeolocate} for handling IP addresses, and **rgeolocate** \citep{rgeolocate} for direct IP geolocation.
+In this article we review the use cases for particular operations over web data, the limitations in base R when it comes to performing those operations, and a suite of R packages designed to overcome them: \CRANpkg{webreadr} \citep{webreadr}, designed for reading access logs in,
+\CRANpkg{urltools} \citep{urltools}, for URL manipulation, \CRANpkg{iptools} \citep{iptools} for handling IP addresses, and \CRANpkg{rgeolocate} \citep{rgeolocate} for direct IP geolocation.
 
@@
 
-To make reading access logs into R as easy as possible we created the **webreadr** package.
+To make reading access logs into R as easy as possible we created the \CRANpkg{webreadr} package.
 This contains user-friendly equivalents to `read.table` for each type of log, detecting the fields that should appear, converting the timestamps into POSIX objects, and merging fields or splitting fields where necessary.
@@
 
-As the name suggests, it is built not on top of base R but on top of the **readr** \citep{readr} package, allowing us to take advantage of substantial speed improvements that package's base functions have over base R.
-These improvements can be seen in the visualisation below, which uses **microbenchmark** \citep{microbenchmark} to compare 100 reads of a 600,000-line "squid" formatted file with webreadr to the same operation performed in base R:
+As the name suggests, it is built not on top of base R but on top of the \CRANpkg{readr} \citep{readr} package, allowing us to take advantage of substantial speed improvements that package's base functions have over base R.
+These improvements can be seen in the visualisation below, which uses \CRANpkg{microbenchmark} \citep{microbenchmark} to compare 100 reads of a 600,000-line "squid" formatted file with webreadr to the same operation performed in base R:
 
@@
 
-To solve this common problem in analysing request logs, the **urltools** [@urltools] package was created. This contains a function,
+To solve this common problem in analysing request logs, the \CRANpkg{urltools} [@urltools] package was created. This contains a function,
 <code>url_decode</code>, which decodes URLs - but relies on vectorised, compiled code.
-Benchmarking the two approaches against each other shows that the `urltools` implementation is approximately 60-70 times faster over large datasets.
-Again using **microbenchmark**, if we compare the vectorised decoding of 1,000,000 URLs with `urltools` against `URLdecode` in a `vapply` loop, we see a 60-70 times speed improvement of urltools over base R:
+Benchmarking the two approaches against each other shows that the \CRANpkg{urltools} implementation is approximately 60-70 times faster over large datasets.
+Again using **microbenchmark**, if we compare the vectorised decoding of 1,000,000 URLs with \CRANpkg{urltools} against `URLdecode` in a `vapply` loop, we see a 60-70 times speed improvement of urltools over base R:
 
@@
 
-Base R does not have native code to parse URLs, but the **httr** package \citep{httr} contains a function, `parse_url`, designed to do just thatm built on R's regular expressions.
+Base R does not have native code to parse URLs, but the \CRANpkg{httr} package \citep{httr} contains a function, `parse_url`, designed to do just thatm built on R's regular expressions.
 This function is not vectorised, does not make use of compiled code internally, and produces a list rather than data.frame, making looping over a set of URLs to parse each one a time-consuming and awkward experience.
@@
 
-`urltools` contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame.
-Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, not having designed to be, does not.
+\CRANpkg{urltools} contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame.
+Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, which was never designed with vectorisation in mind, does not.
 Indeed, benchmarking showed that `url_parse` is approximately 570 times faster than httr's equivalent function:
@@
 ```
+
+As a result of this functionality, urltools provides a package to make URL manipulation faster, easier and far more accessible, reducing the burden associated with munging access logs or similar datasets and allowing a researcher to get to the statistical analysis faster.
+
 ## IP manipulation and geolocation
 
-## Summary
+Access logs also contain IP addresses - unique numeric values that identify a particular computer or network in the context of the internet.
 
-This file is only a basic article template.
-For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](http://journal.r-project.org/latex/RJauthorguide.pdf).
+<!-- Bob, I'll rely on you to fill out bits on how iptools is useful for this kind of work; I tend to primarily rely on rgeolocate since my ip-needs are more limited. -->
+
+As a side-effect of how IP addresses tend to be assigned - in geographic blocks, to individual machines or to local networks - they can be used to geolocate requests, identifying where in the world the request came from, sometimes down to the level of individual post codes or pairs of latitude/longitude coordinates.
+
+This is tremendously useful in industry, where the geographic reach of a service has substantial implications for its viability and survivability, and in academia, where the locality of internet-provided information and the breadth of internet access are active concerns and areas of study \citep{barriers}.
+
+Many services and databases exist for extracting geographic metadata from IP addresses.
+One of the most common is the service provided by MaxMind, which has both proprietary and openly-licensed databases, in binary and comma-separated formats.
+The free databases have been used by various web APIs, which makes the data they contain accessible from R.
+Unfortunately, dependence on web APIs means that handling large numbers of IP addresses can be very slow (they tend to be designed to only accept one IP address at a time, and may contain throttling beyond that) and has privacy concerns, since it essentially means sending user IP addresses to a third party.
+And even without these issues, there are no convenient wrappers for those APIs available on CRAN: users have to hand-roll their own.
+
+With these concerns in mind, we wrote the \CRANpkg{rgeolocate} package. Through \pkg{httr} this contains convenient, vectorised bindings
+to various web services that provide geographic metadata about IP addresses.
+More importantly, using the \CRANpkg{Rcpp} package to integrate C++ and R, \pkg{rgeolocate} also features a direct, compiled binding to the MaxMind API.
+This means that local binary databases can also be queried, which is far faster and more robust than web-based equivalents and avoids the privacy concerns associated with transmitting users' IP addresses externally.
+
+The MaxMind API requires a paid or free binary database - one of which, for country-level IP resolution, is included in \pkg{rgeolocate} - and allows you to retrieve the continent, country name or ISO code, region or city name, tzdata-compatible timezone, longitude, latitude or connection type of a particular IP address.
+Multiple fields can be selected (although which are available depends on the type of database used), and results are returned in a data.frame:
+
+```{r, eval=TRUE}
+
+# Load rgeolocate
+library(rgeolocate)
+
+# Find the rgeolocate-provided binary database
+geo_file <- system.file("extdata","GeoLite2-Country.mmdb", package = "rgeolocate")
+
+# Put together some example IP addresses
+ip_addresses <- c("174.62.175.82", "196.200.60.51")
+
+# Geolocate
+rgeolocate::maxmind(ip_addresses, geo_file, fields = c("continent_name", "country_code"))
+
+```
+
+Direct speed comparisons aren't possible, since the functionality it provides is not (to the authors' knowledge) replicated in
+other R packages, but it is certainly faster (and more secure) than internet-dependent alternatives.
+
+## Conclusions and further work
+
+In this research article we have demonstrated a set of tools for handling access logs during every stage of the data cleaning pipeline - reading them in with \pkg{webreadr}, decoding, manipulating and extracting value from URLs with \pkg{urltools}, and retrieving geographic metadata from IP addresses with \pkg{iptools} and \pkg{rgeolocate}.
+We have also demonstrated the dramatic speed improvements in using these tools in preference to existing methods within R.
+
+Further work - integrating more sources of geolocation information within \pkg{rgeolocate}, supporting UTF-8 URL decoding within \pkg{urltools}, and increasing IPv6 support in \pkg{iptools} - would make these packages even more useful to Human-Computer Interaction researchers and other specialists who rely on access logs as primary data sources.
+Even without that functionality, however, this suite of packages is a dramatic improvement upon the status quo.
 

### 2015-12-20 paper/paper.tex
FIRST DRAFT DONE

@@
 \title{R Packages to Aid in Handling Web Access Logs}
-\author{by Oliver Keyes, Bob Rudis}
+\author{by Oliver Keyes, Bob Rudis, Jay Jacobs}
 
@@
 the internet - explorations commonly performed in R. In this paper we
-explain and demonstrate a series of packages - \textbf{webreadr},
-\textbf{urltools}, \textbf{iptools} and \textbf{rgeolocate} - designed
-to efficiently read in, parse and munge access log data, allowing
-researchers to handle it easily.
+explain and demonstrate a series of packages - \CRANpkg{webreadr},
+\CRANpkg{urltools}, \CRANpkg{iptools} and \CRANpkg{rgeolocate} -
+designed to efficiently read in, parse and munge access log data,
+allowing researchers to handle it easily.
 }
@@
 functions, and additional functionality is made available for handling
-web-accessible data through packages such as \textbf{httr} \citep{httr}.
+web-accessible data through packages such as \CRANpkg{httr}
+\citep{httr}.
 
@@
 base R when it comes to performing those operations, and a suite of R
-packages designed to overcome them: \textbf{webreadr} \citep{webreadr},
-designed for reading access logs in, \textbf{urltools} \citep{urltools},
-for URL manipulation, \textbf{iptools} \citep{rgeolocate} for handling
-IP addresses, and \textbf{rgeolocate} \citep{rgeolocate} for direct IP
-geolocation.
+packages designed to overcome them: \CRANpkg{webreadr} \citep{webreadr},
+designed for reading access logs in, \CRANpkg{urltools}
+\citep{urltools}, for URL manipulation, \CRANpkg{iptools}
+\citep{iptools} for handling IP addresses, and \CRANpkg{rgeolocate}
+\citep{rgeolocate} for direct IP geolocation.
 
@@
 To make reading access logs into R as easy as possible we created the
-\textbf{webreadr} package. This contains user-friendly equivalents to
+\CRANpkg{webreadr} package. This contains user-friendly equivalents to
 \texttt{read.table} for each type of log, detecting the fields that
@@
 As the name suggests, it is built not on top of base R but on top of the
-\textbf{readr} \citep{readr} package, allowing us to take advantage of
+\CRANpkg{readr} \citep{readr} package, allowing us to take advantage of
 substantial speed improvements that package's base functions have over
 base R. These improvements can be seen in the visualisation below, which
-uses \textbf{microbenchmark} \citep{microbenchmark} to compare 100 reads
-of a 600,000-line ``squid'' formatted file with webreadr to the same
-operation performed in base R:
+uses \CRANpkg{microbenchmark} \citep{microbenchmark} to compare 100
+reads of a 600,000-line ``squid'' formatted file with webreadr to the
+same operation performed in base R:
 
@@
 To solve this common problem in analysing request logs, the
-\textbf{urltools} {[}@urltools{]} package was created. This contains a
+\CRANpkg{urltools} {[}@urltools{]} package was created. This contains a
 function, url\_decode, which decodes URLs - but relies on vectorised,
 compiled code. Benchmarking the two approaches against each other shows
-that the \texttt{urltools} implementation is approximately 60-70 times
+that the \CRANpkg{urltools} implementation is approximately 60-70 times
 faster over large datasets. Again using \textbf{microbenchmark}, if we
-compare the vectorised decoding of 1,000,000 URLs with \texttt{urltools}
-against \texttt{URLdecode} in a \texttt{vapply} loop, we see a 60-70
-times speed improvement of urltools over base R:
+compare the vectorised decoding of 1,000,000 URLs with
+\CRANpkg{urltools} against \texttt{URLdecode} in a \texttt{vapply} loop,
+we see a 60-70 times speed improvement of urltools over base R:
 
@@
 
-Base R does not have native code to parse URLs, but the \textbf{httr}
+Base R does not have native code to parse URLs, but the \CRANpkg{httr}
 package \citep{httr} contains a function, \texttt{parse\_url}, designed
@@
 
-\texttt{urltools} contains \texttt{url\_parse} - which does the same
+\CRANpkg{urltools} contains \texttt{url\_parse} - which does the same
 thing as the equivalent \texttt{httr} functionality, but in a vectorised
@@
 context of parsing and processing access logs, this is far more useful,
-because it works efficiently over large sets: httr's functionality, not
-having designed to be, does not. Indeed, benchmarking showed that
-\texttt{url\_parse} is approximately 570 times faster than httr's
-equivalent function:
+because it works efficiently over large sets: httr's functionality,
+which was never designed with vectorisation in mind, does not. Indeed,
+benchmarking showed that \texttt{url\_parse} is approximately 570 times
+faster than httr's equivalent function:
 
@@
 influenced by the style of the \emph{lubridate} package
-\citep{lubridate}, \texttt{urltools} contains functions to get or set
-\emph{individual} components:
+\citep{lubridate}, \texttt{urltools} contains functions to get or
+\emph{set} individual components:
 
@@
 \begin{Sinput}
+# Load urltools and construct a URL
 library(urltools)
 url <- "http://www.google.com/"
+
+# Get the scheme
 scheme(url)
@@
 \begin{Sinput}
+# Set the scheme and observe the modification to the resulting URL
 scheme(url) <- "https"
@@
 
+As a result of this functionality, urltools provides a package to make
+URL manipulation faster, easier and far more accessible, reducing the
+burden associated with munging access logs or similar datasets and
+allowing a researcher to get to the statistical analysis faster.
+
 \subsection{IP manipulation and
 geolocation}\label{ip-manipulation-and-geolocation}
+
+Access logs also contain IP addresses - unique numeric values that
+identify a particular computer or network in the context of the
+internet.
+
+As a side-effect of how IP addresses tend to be assigned - in geographic
+blocks, to individual machines or to local networks - they can be used
+to geolocate requests, identifying where in the world the request came
+from, sometimes down to the level of individual post codes or pairs of
+latitude/longitude coordinates.
+
+This is tremendously useful in industry, where the geographic reach of a
+service has substantial implications for its viability and
+survivability, and in academia, where the locality of internet-provided
+information and the breadth of internet access are active concerns and
+areas of study \citep{barriers}.
+
+Many services and databases exist for extracting geographic metadata
+from IP addresses. One of the most common is the service provided by
+MaxMind, which has both proprietary and openly-licensed databases, in
+binary and comma-separated formats. The free databases have been used by
+various web APIs, which makes the data they contain accessible from R.
+Unfortunately, dependence on web APIs means that handling large numbers
+of IP addresses can be very slow (they tend to be designed to only
+accept one IP address at a time, and may contain throttling beyond that)
+and has privacy concerns, since it essentially means sending user IP
+addresses to a third party. And even without these issues, there are no
+convenient wrappers for those APIs available on CRAN: users have to
+hand-roll their own.
+
+With these concerns in mind, we wrote the \CRANpkg{rgeolocate} package.
+Through \pkg{httr} this contains convenient, vectorised bindings to
+various web services that provide geographic metadata about IP
+addresses. More importantly, using the \CRANpkg{Rcpp} package to
+integrate C++ and R, \pkg{rgeolocate} also features a direct, compiled
+binding to the MaxMind API. This means that local binary databases can
+also be queried, which is far faster and more robust than web-based
+equivalents and avoids the privacy concerns associated with transmitting
+users' IP addresses externally.
+
+The MaxMind API requires a paid or free binary database - one of which,
+for country-level IP resolution, is included in \pkg{rgeolocate} - and
+allows you to retrieve the continent, country name or ISO code, region
+or city name, tzdata-compatible timezone, longitude, latitude or
+connection type of a particular IP address. Multiple fields can be
+selected (although which are available depends on the type of database
+used), and results are returned in a data.frame:
+
+\begin{Schunk}
+\begin{Sinput}
+# Load rgeolocate
+library(rgeolocate)
+
+# Find the rgeolocate-provided binary database
+geo_file <- system.file("extdata","GeoLite2-Country.mmdb", package = "rgeolocate")
+
+# Put together some example IP addresses
+ip_addresses <- c("174.62.175.82", "196.200.60.51")
+
+# Geolocate
+rgeolocate::maxmind(ip_addresses, geo_file, fields = c("continent_name", "country_code"))
+\end{Sinput}
+\begin{Soutput}
+#>   continent_name country_code
+#> 1  North America           US
+#> 2         Africa           ML
+\end{Soutput}
+\end{Schunk}
 
@@
 \href{mailto:author2@work}{\nolinkurl{author2@work}}
+
+\address{
+Jay Jacobs\\
+Magical Python Land\\
+line 1\\ line 2\\
+}
+\href{mailto:author3@work}{\nolinkurl{author3@work}}

### 2015-12-21 urltools_benchmarks.R
Suggesting some image improvements

@@ -26,7 +26,7 @@ decoding_benchmarks <- microbenchmark({(urltools_decode(example_urls))},
 levels(decoding_benchmarks$expr)[grepl(x = levels(decoding_benchmarks$expr), pattern = "urltools")] <- "urltools"
 levels(decoding_benchmarks$expr)[levels(decoding_benchmarks$expr) != "urltools"] <- "base R"
 
-# Plot
+# Original Plot
 ggsave(file = "./paper/decoding_benchmarks.png",
        plot = autoplot(decoding_benchmarks) + theme_fivethirtynine(base_size = 14) +
          labs(title = "Decoding 1m URLs, base R versus urltools (log-scaled)"))
@@ -54,3 +54,46 @@ ggsave(file = "./paper/parsing_benchmarks.png",
 # Save the results to disc, because nobody wants to lose benchmarks that take _this long_ to compute.
 save(decoding_benchmarks, parsing_benchmarks, file = "urltools_benchmarks.RData")
 
+### Jay doing some updates
+load(file = "urltools_benchmarks.RData")
+
+dcode <- as.data.frame(decoding_benchmarks)
+dcode$time.sec <- dcode$time/1e+9
+
+mbrk <- c(seq(2,9), seq(20,90,by=10))
+mbrk.df <- data.frame(y=mbrk, expr=NA, time.sec=NA)
+brk <- c(1, 10, 100, 200)
+gg <- ggplot(dcode, aes(expr, time.sec)) + 
+  geom_hline(data=mbrk.df, aes(yintercept=y), color="gray95") +
+  geom_violin(fill="steelblue", color="steelblue") +
+  scale_y_log10("Time [seconds]", breaks=brk, limits=c(1,200), expand=c(0,0)) +
+  coord_flip() +
+  ggtitle("Decoding 1m URLs, base R versus urltools (log-scaled)") +
+  theme(panel.background=element_rect(fill=NA, color="gray75"),
+        panel.grid.major=element_line(color="gray75"),
+        panel.grid.minor=element_blank(),
+        axis.title.y=element_blank(),
+        axis.ticks=element_blank())
+ggsave(gg, file = "./paper/decoding_benchmarks-jay.png")
+       
+
+## parsing_benchmarks
+pcode <- as.data.frame(parsing_benchmarks)
+pcode$time.sec <- pcode$time/1e+9
+
[…]

### 2015-12-21 webreadr_benchmarks.R

@@ -36,3 +36,23 @@ ggsave(file = "./paper/reading_benchmarks.png",
 # Save benchmarks
 save(reading_benchmarks, file = "webreadr_benchmarks.RData")
 
+### Updates from Jay
+
+load(file="webreadr_benchmarks.RData")
+## reading_benchmarks
+rdcode <- as.data.frame(reading_benchmarks)
+rdcode$time.sec <- rdcode$time/1e+9
+
+gg <- ggplot(rdcode, aes(expr, time.sec)) + 
+  #geom_hline(data=mbrk.df, aes(yintercept=y), color="gray95") +
+  geom_violin(fill="steelblue", color="steelblue") +
+  scale_y_continuous("Time [seconds]", breaks=seq(1,6), limits=c(1,6), expand=c(0,0)) +
+  coord_flip() +
+  ggtitle("Reading 600k lines of access log, base R versus webreadr") +
+  theme(panel.background=element_rect(fill=NA, color="gray75"),
+        panel.grid.major=element_line(color="gray75"),
+        #panel.grid.minor=element_blank(),
+        axis.title.y=element_blank(),
+        axis.ticks=element_blank())
+gg
+ggsave(gg, file = "./paper/reading_benchmarks-jay.png")

### 2015-12-26 paper/paper.Rmd
copyedit and style

@@
   of human behaviour on the internet - explorations commonly performed in R. In this paper we explain and demonstrate
-  a series of packages - \CRANpkg{webreadr}, \CRANpkg{urltools}, \CRANpkg{iptools} and \CRANpkg{rgeolocate} - designed to efficiently read in, parse
+  a series of packages designed to efficiently read in, parse
   and munge access log data, allowing researchers to handle it easily.
@@
 
-To make reading access logs into R as easy as possible we created the \CRANpkg{webreadr} package.
-This contains user-friendly equivalents to `read.table` for each type of log, detecting the fields that should appear, converting the timestamps into POSIX objects, and merging fields or splitting fields where necessary.
-The package contains four core functions, one for each form of log: `read_clf` for the Common Log Format, `read_combined` for the Combined Log Format, and `read_squid` and `read_aws` for Squid and AWS formats respectively.
+To make reading access logs into R as easy as possible we created the \pkg{webreadr} package.
+This contains user-friendly equivalents to \code{read.table} for each type of log, detecting the fields that should appear, converting the timestamps into POSIX objects, and merging fields or splitting fields where necessary.
+The package contains four core functions, one for each form of log: \code{read\_clf} for the Common Log Format, \code{read\_combined} for the Combined Log Format, and \code{read\_squid} and \code{read\_aws} for Squid and AWS formats respectively.
 Each one abstracts away the complexity of specifying column names and formats, and instead allows a researcher to read a file in with a minimal amount of work.
@@
 
-**webreadr** takes 1.2 seconds to read this file in at a minimum, and 1.48 at maximum, with a median of 1.33.
+\pkg{webreadr} takes 1.2 seconds to read this file in at a minimum, and 1.48 at maximum, with a median of 1.33.
 This is consistently 3.5-6 times faster than the equivalent base R functionality - and is also, as explained, far easier to use.
@@
 Both values can be percent-encoded, allowing them toinclude characters that are valid but reserved by the URL specification as having special meanings ("reserved characters").
-A `#` symbol, for example, is encoded as <code>%23</code>: a percentage symbol, followed by the byte-code for that character.
+A \samp{\#} symbol, for example, is encoded as \samp{\%23}: a percentage symbol, followed by the byte-code for that character.
 
@@
 The solution is to be able to consistently decode URLs, which makes URL-based data far easier to analyse.
-Base R does contain a function, <code>URLdecode</code>, for doing just that, but is neither vectorised nor based on compiled code; with large datasets it takes an extremely long time.
+Base R does contain a function, \code{URLdecode}, for doing just that, but is neither vectorised nor based on compiled code; with large datasets it takes an extremely long time.
 
-To solve this common problem in analysing request logs, the \CRANpkg{urltools} [@urltools] package was created. This contains a function,
-<code>url_decode</code>, which decodes URLs - but relies on vectorised, compiled code.
-Benchmarking the two approaches against each other shows that the \CRANpkg{urltools} implementation is approximately 60-70 times faster over large datasets.
-Again using **microbenchmark**, if we compare the vectorised decoding of 1,000,000 URLs with \CRANpkg{urltools} against `URLdecode` in a `vapply` loop, we see a 60-70 times speed improvement of urltools over base R:
+To solve this common problem in analysing request logs, the \pkg{urltools} package was created. This contains a function,
+\code{url\_decode}, which decodes URLs - but relies on vectorised, compiled code.
+Benchmarking the two approaches against each other shows that the \pkg{urltools} implementation is approximately 60-70 times faster over large datasets.
+Again using \pkg{microbenchmark}, if we compare the vectorised decoding of 1,000,000 URLs with \pkg{urltools} against \code{URLdecode} in a \code{vapply} loop, we see a 60-70 times speed improvement:
 
@@
     \includegraphics[scale=0.4]{decoding_benchmarks}
-    \caption{Results of microbenchmark run: url\char`_decode versus base-R equivalent code}
+    \caption{Results of microbenchmark run: \code{url\_decode} versus base-R equivalent code}
 \end{figure}

### 2015-12-26 paper/paper.tex
copyedit and style

@@
 
-\abstract{
+\abstract{%
 Web access logs contain information on HTTP(S) requests and form a key
@@
 the internet - explorations commonly performed in R. In this paper we
-explain and demonstrate a series of packages - \CRANpkg{webreadr},
-\CRANpkg{urltools}, \CRANpkg{iptools} and \CRANpkg{rgeolocate} -
-designed to efficiently read in, parse and munge access log data,
-allowing researchers to handle it easily.
+explain and demonstrate a series of packages designed to efficiently
+read in, parse and munge access log data, allowing researchers to handle
+it easily.
 }
@@
 To make reading access logs into R as easy as possible we created the
-\CRANpkg{webreadr} package. This contains user-friendly equivalents to
-\texttt{read.table} for each type of log, detecting the fields that
-should appear, converting the timestamps into POSIX objects, and merging
-fields or splitting fields where necessary. The package contains four
-core functions, one for each form of log: \texttt{read\_clf} for the
-Common Log Format, \texttt{read\_combined} for the Combined Log Format,
-and \texttt{read\_squid} and \texttt{read\_aws} for Squid and AWS
-formats respectively. Each one abstracts away the complexity of
-specifying column names and formats, and instead allows a researcher to
-read a file in with a minimal amount of work.
+\pkg{webreadr} package. This contains user-friendly equivalents to
+\code{read.table} for each type of log, detecting the fields that should
+appear, converting the timestamps into POSIX objects, and merging fields
+or splitting fields where necessary. The package contains four core
+functions, one for each form of log: \code{read\_clf} for the Common Log
+Format, \code{read\_combined} for the Combined Log Format, and
+\code{read\_squid} and \code{read\_aws} for Squid and AWS formats
+respectively. Each one abstracts away the complexity of specifying
+column names and formats, and instead allows a researcher to read a file
+in with a minimal amount of work.
 
@@
 
-\textbf{webreadr} takes 1.2 seconds to read this file in at a minimum,
-and 1.48 at maximum, with a median of 1.33. This is consistently 3.5-6
-times faster than the equivalent base R functionality - and is also, as
+\pkg{webreadr} takes 1.2 seconds to read this file in at a minimum, and
+1.48 at maximum, with a median of 1.33. This is consistently 3.5-6 times
+faster than the equivalent base R functionality - and is also, as
 explained, far easier to use.
@@
 that are valid but reserved by the URL specification as having special
-meanings (``reserved characters''). A \texttt{\#} symbol, for example,
-is encoded as \%23: a percentage symbol, followed by the byte-code for
-that character.
+meanings (``reserved characters''). A \samp{\#} symbol, for example, is
+encoded as \samp{\%23}: a percentage symbol, followed by the byte-code
+for that character.
 
@@
 URL-based data far easier to analyse. Base R does contain a function,
-URLdecode, for doing just that, but is neither vectorised nor based on
-compiled code; with large datasets it takes an extremely long time.
+\code{URLdecode}, for doing just that, but is neither vectorised nor
+based on compiled code; with large datasets it takes an extremely long
+time.
 
 To solve this common problem in analysing request logs, the
-\CRANpkg{urltools} {[}@urltools{]} package was created. This contains a
-function, url\_decode, which decodes URLs - but relies on vectorised,
+\pkg{urltools} package was created. This contains a function,
+\code{url\_decode}, which decodes URLs - but relies on vectorised,
 compiled code. Benchmarking the two approaches against each other shows
-that the \CRANpkg{urltools} implementation is approximately 60-70 times
-faster over large datasets. Again using \textbf{microbenchmark}, if we
-compare the vectorised decoding of 1,000,000 URLs with
-\CRANpkg{urltools} against \texttt{URLdecode} in a \texttt{vapply} loop,
-we see a 60-70 times speed improvement of urltools over base R:
+that the \pkg{urltools} implementation is approximately 60-70 times
+faster over large datasets. Again using \pkg{microbenchmark}, if we
+compare the vectorised decoding of 1,000,000 URLs with \pkg{urltools}
+against \code{URLdecode} in a \code{vapply} loop, we see a 60-70 times
+speed improvement:
 
@@
 
-\subsection{Summary}\label{summary}
+Direct speed comparisons aren't possible, since the functionality it
+provides is not (to the authors' knowledge) replicated in other R
+packages, but it is certainly faster (and more secure) than
+internet-dependent alternatives.
 
-This file is only a basic article template. For full details of
-\emph{The R Journal} style and information on how to prepare your
-article for submission, see the
-\href{http://journal.r-project.org/latex/RJauthorguide.pdf}{Instructions
-for Authors}.
+\subsection{Conclusions and further
+work}\label{conclusions-and-further-work}
+
+In this research article we have demonstrated a set of tools for
+handling access logs during every stage of the data cleaning pipeline -
+reading them in with \pkg{webreadr}, decoding, manipulating and
+extracting value from URLs with \pkg{urltools}, and retrieving
+geographic metadata from IP addresses with \pkg{iptools} and
+\pkg{rgeolocate}. We have also demonstrated the dramatic speed
+improvements in using these tools in preference to existing methods
+within R.
+
+Further work - integrating more sources of geolocation information
+within \pkg{rgeolocate}, supporting UTF-8 URL decoding within
+\pkg{urltools}, and increasing IPv6 support in \pkg{iptools} - would
+make these packages even more useful to Human-Computer Interaction
+researchers and other specialists who rely on access logs as primary
+data sources. Even without that functionality, however, this suite of
+packages is a dramatic improvement upon the status quo.
 
@@
 
-\address{
+\address{%
 Oliver Keyes\\
@@
 
-\address{
+\address{%
 Bob Rudis\\
@@
 
-\address{
+\address{%
 Jay Jacobs\\

### 2015-12-26 paper/paper.Rmd
formatting

@@
 
-The standard for URLs \citep{RFC1738} divides them into a heirarchical sequence of components - the scheme ("http"), host ("en.wikipedia.org"), port ("800"), path ("wiki/Main_Page") and searchpart, or parameters ("action=edit").
-Together, these make
-up a URL ("http://en.wikipedia.org:800/wiki/Main_Page?action=edit").
+The standard for URLs \citep{RFC1738} divides them into a heirarchical sequence of components - the scheme (\samp{http}), host (\samp{en.wikipedia.org}), port (\samp{800}), path (\samp{wiki/Main\_Page}) and searchpart, or parameters (\samp{action=edit}).
+Together, these make up a URL
+\linebreak (\samp{http://en.wikipedia.org:800/wiki/Main\_Page?action=edit}).
 

### 2015-12-26 paper/paper.tex
formatting

@@
 The standard for URLs \citep{RFC1738} divides them into a heirarchical
-sequence of components - the scheme (``http''), host
-(``en.wikipedia.org''), port (``800''), path (``wiki/Main\_Page'') and
-searchpart, or parameters (``action=edit''). Together, these make up a
-URL (``\url{http://en.wikipedia.org:800/wiki/Main_Page?action=edit}'').
+sequence of components - the scheme (\samp{http}), host
+(\samp{en.wikipedia.org}), port (\samp{800}), path
+(\samp{wiki/Main\_Page}) and searchpart, or parameters
+(\samp{action=edit}). Together, these make up a URL
+\linebreak (\samp{http://en.wikipedia.org:800/wiki/Main\_Page?action=edit}).
 

### 2015-12-26 paper/paper.Rmd
cleanup, .gitignore improvements

@@
 
-Base R does not have native code to parse URLs, but the \CRANpkg{httr} package \citep{httr} contains a function, `parse_url`, designed to do just thatm built on R's regular expressions.
-This function is not vectorised, does not make use of compiled code internally, and produces a list rather than data.frame, making looping over a set of URLs to parse each one a time-consuming and awkward experience.
+Base R does not have native code to parse URLs, but the \CRANpkg{httr} package \citep{httr} contains a function, \code{parse\_url}, designed to do just that.
+Built on R's regular expressions, this function is not vectorised, does not make use of compiled code internally, and produces a list rather than data.frame, making looping over a set of URLs to parse each one a time-consuming and awkward experience.
 This is understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse URLs *en masse*.
 
-\CRANpkg{urltools} contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame.
+\pkg{urltools} contains \code{url_parse} - which does the same thing as the equivalent \pkg{httr} functionality, but in a vectorised way, relying on compiled code, and producing a data.frame.
 Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, which was never designed with vectorisation in mind, does not.
-Indeed, benchmarking showed that `url_parse` is approximately 570 times faster than httr's equivalent function:
+Indeed, benchmarking showed that \code{url\_parse} is approximately 570 times faster than httr's equivalent function:
 
@@
     \includegraphics[scale=0.4]{parsing_benchmarks}
-    \caption{Results of microbenchmark run: url\char`_parse versus httr's equivalent code}
+    \caption{Results of microbenchmark run: \code{url\_parse} versus httr's equivalent code}
 \end{figure}

### 2015-12-26 paper/paper.tex
cleanup, .gitignore improvements

@@
 Base R does not have native code to parse URLs, but the \CRANpkg{httr}
-package \citep{httr} contains a function, \texttt{parse\_url}, designed
-to do just thatm built on R's regular expressions. This function is not
+package \citep{httr} contains a function, \code{parse\_url}, designed to
+do just that. Built on R's regular expressions, this function is not
 vectorised, does not make use of compiled code internally, and produces
@@
 
-\CRANpkg{urltools} contains \texttt{url\_parse} - which does the same
-thing as the equivalent \texttt{httr} functionality, but in a vectorised
-way, relying on compiled code, and producing a data.frame. Within the
-context of parsing and processing access logs, this is far more useful,
-because it works efficiently over large sets: httr's functionality,
-which was never designed with vectorisation in mind, does not. Indeed,
-benchmarking showed that \texttt{url\_parse} is approximately 570 times
+\pkg{urltools} contains \code{url_parse} - which does the same thing as
+the equivalent \pkg{httr} functionality, but in a vectorised way,
+relying on compiled code, and producing a data.frame. Within the context
+of parsing and processing access logs, this is far more useful, because
+it works efficiently over large sets: httr's functionality, which was
+never designed with vectorisation in mind, does not. Indeed,
+benchmarking showed that \code{url\_parse} is approximately 570 times
 faster than httr's equivalent function:

### 2015-12-26 paper/paper.Rmd
copyedit, improve

@@
 
-A vector of URLs passed into `url_parse` produces a data.frame, with one column for each of the IETF-supported components, and empty strings representing components that could not be found.
-Additionally, influenced by the style of the *lubridate* package \citep{lubridate}, `urltools` contains functions to get or *set* individual components:
+A vector of URLs passed into \code{url\_parse} produces a data.frame, with one column for each of the IETF-supported components, and empty strings representing components that could not be found.
+Additionally, influenced by the style of the \CRANpkg{lubridate} package \citep{lubridate}, \pkg{urltools} contains functions to get or *set* individual components:
 
@@
 
-As a result of this functionality, urltools provides a package to make URL manipulation faster, easier and far more accessible, reducing the burden associated with munging access logs or similar datasets and allowing a researcher to get to the statistical analysis faster.
+As a result of this functionality, \pkg{urltools} makes URL manipulation faster, easier and far more accessible, reducing the burden associated with munging access logs or similar datasets and allowing a researcher to get to the statistical analysis faster.
 
@@
 
-With these concerns in mind, we wrote the \CRANpkg{rgeolocate} package. Through \pkg{httr} this contains convenient, vectorised bindings
+With these concerns in mind, we wrote the \pkg{rgeolocate} package. Through \pkg{httr} this contains convenient, vectorised bindings
 to various web services that provide geographic metadata about IP addresses.

### 2015-12-26 paper/paper.Rmd
update

@@
 
-\pkg{urltools} contains \code{url_parse} - which does the same thing as the equivalent \pkg{httr} functionality, but in a vectorised way, relying on compiled code, and producing a data.frame.
+\pkg{urltools} contains \code{url\_parse} - which does the same thing as the equivalent \pkg{httr} functionality, but in a vectorised way, relying on compiled code, and producing a data.frame.
 Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, which was never designed with vectorisation in mind, does not.

### 2015-12-26 paper/paper.tex
update

@@
 
-\pkg{urltools} contains \code{url_parse} - which does the same thing as
+\pkg{urltools} contains \code{url\_parse} - which does the same thing as
 the equivalent \pkg{httr} functionality, but in a vectorised way,
@@
 
-A vector of URLs passed into \texttt{url\_parse} produces a data.frame,
+A vector of URLs passed into \code{url\_parse} produces a data.frame,
 with one column for each of the IETF-supported components, and empty
 strings representing components that could not be found. Additionally,
-influenced by the style of the \emph{lubridate} package
-\citep{lubridate}, \texttt{urltools} contains functions to get or
+influenced by the style of the \CRANpkg{lubridate} package
+\citep{lubridate}, \pkg{urltools} contains functions to get or
 \emph{set} individual components:
@@
 
-As a result of this functionality, urltools provides a package to make
-URL manipulation faster, easier and far more accessible, reducing the
-burden associated with munging access logs or similar datasets and
-allowing a researcher to get to the statistical analysis faster.
+As a result of this functionality, \pkg{urltools} makes URL manipulation
+faster, easier and far more accessible, reducing the burden associated
+with munging access logs or similar datasets and allowing a researcher
+to get to the statistical analysis faster.
 
@@
 
-With these concerns in mind, we wrote the \CRANpkg{rgeolocate} package.
+With these concerns in mind, we wrote the \pkg{rgeolocate} package.
 Through \pkg{httr} this contains convenient, vectorised bindings to

### 2015-12-26 paper/paper.Rmd
start iptools section

@@
 
+Working with these values allows an analyst to validate their data (by checking for false or spoof addresses) and is a necessary prerequisite to the use of some IP geolocation systems (covered later) which require the numeric forms of IP addresses (\samp{3232235521}) rather than the standard \dfn{dotted-decimal} form (\samp{192.168.0.1}) found in access logs.
+
+Based around the \emph{Boost ASIO} C++ library \citep{boostcite}, \pkg{iptools} is a package designed for this kind of IP manipulation (and more).
+Designed around combined code, it is extremely fast (for that code that does not rely on internet connections) and boasts a range of features.
+
+One of the most crucial is \code{ip\_classify}, which, when provided with a vector of IP addresses, identifies whether they follow the IPv4 or IPv6 standard.
+As a side-effect of this, it can be used to identify if IP addresses are invalid, or spoofed, prior to further work based on the assumption that they are correct:
+
+```{r, eval=TRUE}
+# Load iptools and construct a vector of IPs
+library(iptools)
+ip_addresses <- c("192.168.0.1", "2607:f8b0:4006:80b::1004", "Chewie")
+ip_classify(ip_addresses)
+
+```
+
 <!-- Bob, I'll rely on you to fill out bits on how iptools is useful for this kind of work; I tend to primarily rely on rgeolocate since my ip-needs are more limited. -->

### 2015-12-26 paper/paper.tex
start iptools section

@@
 
+Working with these values allows an analyst to validate their data (by
+checking for false or spoof addresses) and is a necessary prerequisite
+to the use of some IP geolocation systems (covered later) which require
+the numeric forms of IP addresses (\samp{3232235521}) rather than the
+standard \dfn{dotted-decimal} form (\samp{192.168.0.1}) found in access
+logs.
+
+Based around the \emph{Boost ASIO} C++ library \citep{boostcite},
+\pkg{iptools} is a package designed for this kind of IP manipulation
+(and more). Designed around combined code, it is extremely fast (for
+that code that does not rely on internet connections) and boasts a range
+of features.
+
+One of the most crucial is \code{ip\_classify}, which, when provided
+with a vector of IP addresses, identifies whether they follow the IPv4
+or IPv6 standard. As a side-effect of this, it can be used to identify
+if IP addresses are invalid, or spoofed, prior to further work based on
+the assumption that they are correct:
+
+\begin{Schunk}
+\begin{Sinput}
+# Load iptools and construct a vector of IPs
+library(iptools)
+ip_addresses <- c("192.168.0.1", "2607:f8b0:4006:80b::1004", "Chewie")
+ip_classify(ip_addresses)
+\end{Sinput}
+\begin{Soutput}
+#> [1] "IPv4"    "IPv6"    "Invalid"
+\end{Soutput}
+\end{Schunk}
+
 As a side-effect of how IP addresses tend to be assigned - in geographic

### 2015-12-27 paper/paper.Rmd
expand iptools description

@@
 
-## IP manipulation and geolocation
+## IP manipulation
 
 Access logs also contain IP addresses - unique numeric values that identify a particular computer or network in the context of the internet.
-
-Working with these values allows an analyst to validate their data (by checking for false or spoof addresses) and is a necessary prerequisite to the use of some IP geolocation systems (covered later) which require the numeric forms of IP addresses (\samp{3232235521}) rather than the standard \dfn{dotted-decimal} form (\samp{192.168.0.1}) found in access logs.
+Working with these values allows an analyst to validate their data (by checking for false or spoof addresses) and is a necessary prerequisite to the use of some IP geolocation systems (covered later), and extract a limited amount of metadata from the values themselves.
 
 Based around the \emph{Boost ASIO} C++ library \citep{boostcite}, \pkg{iptools} is a package designed for this kind of IP manipulation (and more).
-Designed around combined code, it is extremely fast (for that code that does not rely on internet connections) and boasts a range of features.
+Built around combined code, it is extremely fast (for that code that does not rely on internet connections) and boasts a range of features.
 
@@
 
-<!-- Bob, I'll rely on you to fill out bits on how iptools is useful for this kind of work; I tend to primarily rely on rgeolocate since my ip-needs are more limited. -->
+\pkg{iptools} also contains (for the same purpose) code to identify the actual client's IP address.
+Access requests usually contain not only an IP address but a \dfn{X-Forwarded-For} field - a field identifying which \emph{other} IP addresses the request passed through, if the user who made the request is using some kind of proxy.
+If a user \emph{has} used a proxy, the contents of the IP address field won't actually be them - it will be the last proxy the request went through before getting to the server logging the requests.
+The actual IP address of the user will instead be the earliest value of the X-Forwarded-For field.
 
+The solution is to be able to identify the 'real' IP address, by checking:
+
+1. Whether the X-Forwarded-For field contains any values;
+2. Extracting the earliest non-invalid IP address in that field's values if so, and the contents of the IP address field if not.
+
+With the \code{xff\_extract} function, you can do just that:
+
+```{r, eval = TRUE}
+ip_address <- "192.168.0.1"
+x_forwarded_for <- "foo, 193.168.0.1, 230.98.107.1"
+xff_extract(ip_address, x_forwarded_for)
+```
+
+This returns the IP address field value if X-Forwarded-For is empty, and otherwise splits the X-Forwarded-For field and returns the earliest valid IP address.
+It is fully vectorised and highly useful for analysis predicated on IP addresses being valid, such as geolocation - without this kind of resolution, what you're actually geolocating might be your own servers.
+
+Other operations supported by \pkg{iptools} include the conversion of IP addresses from their standard \dfn{dotted-decimal} form to a numeric form, the extraction of IP ranges (and their contents), resolving IP addresses to hostnames, and a series of datasets covering the IPv4 registry and port database.
+
+The limitation of the package is that it does not yet work on Windows - building ASIO on that platform is a uniquely frustrating experience - and that some operations do not yet support IPv6 (since they require the storage of numbers bigger than R can currently handle).
+
+## Geolocation
 As a side-effect of how IP addresses tend to be assigned - in geographic blocks, to individual machines or to local networks - they can be used to geolocate requests, identifying where in the world the request came from, sometimes down to the level of individual post codes or pairs of latitude/longitude coordinates.

### 2015-12-27 paper/paper.tex
expand iptools description

@@
 
-\subsection{IP manipulation and
-geolocation}\label{ip-manipulation-and-geolocation}
+\subsection{IP manipulation}\label{ip-manipulation}
 
@@
 identify a particular computer or network in the context of the
-internet.
-
-Working with these values allows an analyst to validate their data (by
-checking for false or spoof addresses) and is a necessary prerequisite
-to the use of some IP geolocation systems (covered later) which require
-the numeric forms of IP addresses (\samp{3232235521}) rather than the
-standard \dfn{dotted-decimal} form (\samp{192.168.0.1}) found in access
-logs.
+internet. Working with these values allows an analyst to validate their
+data (by checking for false or spoof addresses) and is a necessary
+prerequisite to the use of some IP geolocation systems (covered later),
+and extract a limited amount of metadata from the values themselves.
 
@@
 \pkg{iptools} is a package designed for this kind of IP manipulation
-(and more). Designed around combined code, it is extremely fast (for
-that code that does not rely on internet connections) and boasts a range
-of features.
+(and more). Built around combined code, it is extremely fast (for that
+code that does not rely on internet connections) and boasts a range of
+features.
 
@@
 \end{Schunk}
+
+\pkg{iptools} also contains (for the same purpose) code to identify the
+actual client's IP address. Access requests usually contain not only an
+IP address but a \dfn{X-Forwarded-For} field - a field identifying which
+\emph{other} IP addresses the request passed through, if the user who
+made the request is using some kind of proxy. If a user \emph{has} used
+a proxy, the contents of the IP address field won't actually be them -
+it will be the last proxy the request went through before getting to the
+server logging the requests. The actual IP address of the user will
+instead be the earliest value of the X-Forwarded-For field.
+
+The solution is to be able to identify the `real' IP address, by
+checking:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Whether the X-Forwarded-For field contains any values;
+\item
+  Extracting the earliest non-invalid IP address in that field's values
+  if so, and the contents of the IP address field if not.
+\end{enumerate}
+
+With the \code{xff\_extract} function, you can do just that:
+
+\begin{Schunk}
+\begin{Sinput}
+ip_address <- "192.168.0.1"
+x_forwarded_for <- "foo, 193.168.0.1, 230.98.107.1"
+xff_extract(ip_address, x_forwarded_for)
+\end{Sinput}
+\begin{Soutput}
+#> [1] "192.168.0.1"
+\end{Soutput}
+\end{Schunk}
+
+This returns the IP address field value if X-Forwarded-For is empty, and
+otherwise splits the X-Forwarded-For field and returns the earliest
+valid IP address. It is fully vectorised and highly useful for analysis
+predicated on IP addresses being valid, such as geolocation - without
+this kind of resolution, what you're actually geolocating might be your
+own servers.
+
+Other operations supported by \pkg{iptools} include the conversion of IP
+addresses from their standard \dfn{dotted-decimal} form to a numeric
+form, the extraction of IP ranges (and their contents), resolving IP
+addresses to hostnames, and a series of datasets covering the IPv4
+registry and port database.
+
+The limitation of the package is that it does not yet work on Windows -
+building ASIO on that platform is a uniquely frustrating experience -
+and that some operations do not yet support IPv6 (since they require the
+storage of numbers bigger than R can currently handle).
+
+\subsection{Geolocation}\label{geolocation}
 

### 2015-12-27 paper/paper.Rmd
add address

@@
     address:
-    - line 1
-    - line 2
-    email:  author1@work
+    - 149 New Montgomery Street, 6th floor
+    - San Francisco, CA, 94105, USA
+    email:  ironholds@gmail.com
   - name: Bob Rudis

### 2015-12-27 paper/paper.Rmd
0.2.0

@@
 The rise of the World Wide Web over the last few decades has made it dramatically easier to access and transfer data, and R \citep{RCore} boasts abundant functionality when it comes to taking data *from* the web.
-Base R itself has simple file downloading and page reading capabilities, through the `download.file` and `readLines` functions, and additional functionality is made available for handling web-accessible data through packages such as \CRANpkg{httr} \citep{httr}.
+Base R itself has simple file downloading and page reading capabilities, through the \code{download.file} and \code{readLines} functions, and additional functionality is made available for handling web-accessible data through packages such as \CRANpkg{httr} \citep{httr}.
 
@@
 As a statistically-oriented programming language, R is commonly used by these same researchers for data analysis, testing and reporting - but it lacks tools designed for the kinds of data sources and formats that are encountered when dealing with access logs.
-In this article we review the use cases for particular operations over web data, the limitations in base R when it comes to performing those operations, and a suite of R packages designed to overcome them: \CRANpkg{webreadr} \citep{webreadr}, designed for reading access logs in,
-\CRANpkg{urltools} \citep{urltools}, for URL manipulation, \CRANpkg{iptools} \citep{iptools} for handling IP addresses, and \CRANpkg{rgeolocate} \citep{rgeolocate} for direct IP geolocation.
+In this article we review the use cases for particular operations over web data, the limitations in base R when it comes to performing those operations, and a suite of R packages designed to overcome them: reading access logs in, manipulating URLs, manipulating IP addresses, and direct IP geolocation.
 
@@
 
-To make reading access logs into R as easy as possible we created the \pkg{webreadr} package.
+To make reading access logs into R as easy as possible we created the \CRANpkg{webreadr} \citep{webreadr} package.
 This contains user-friendly equivalents to \code{read.table} for each type of log, detecting the fields that should appear, converting the timestamps into POSIX objects, and merging fields or splitting fields where necessary.
@@
 
-To solve this common problem in analysing request logs, the \pkg{urltools} package was created. This contains a function,
-\code{url\_decode}, which decodes URLs - but relies on vectorised, compiled code.
+To solve this common problem in analysing request logs, the\CRANpkg{urltools} \citep{urltools} package was created. This contains a function,
+\code{url\_decode}, which decodes URLs and relies on vectorised, compiled code to do so.
 Benchmarking the two approaches against each other shows that the \pkg{urltools} implementation is approximately 60-70 times faster over large datasets.
@@
 
-Based around the \emph{Boost ASIO} C++ library \citep{boostcite}, \pkg{iptools} is a package designed for this kind of IP manipulation (and more).
+Based around the \emph{Boost ASIO} C++ library \citep{boostcite}, \CRANpkg{iptools} \citep{iptools} is a package designed for this kind of IP manipulation (and more).
 Built around combined code, it is extremely fast (for that code that does not rely on internet connections) and boasts a range of features.
@@
 
-With these concerns in mind, we wrote the \pkg{rgeolocate} package. Through \pkg{httr} this contains convenient, vectorised bindings
+With these concerns in mind, we wrote the \CRANpkg{rgeolocate} \citep{rgeolocate} package. Through \pkg{httr} this contains convenient, vectorised bindings
 to various web services that provide geographic metadata about IP addresses.

### 2015-12-27 paper/paper.tex
add address
0.2.0

@@
 the web. Base R itself has simple file downloading and page reading
-capabilities, through the \texttt{download.file} and \texttt{readLines}
+capabilities, through the \code{download.file} and \code{readLines}
 functions, and additional functionality is made available for handling
@@
 base R when it comes to performing those operations, and a suite of R
-packages designed to overcome them: \CRANpkg{webreadr} \citep{webreadr},
-designed for reading access logs in, \CRANpkg{urltools}
-\citep{urltools}, for URL manipulation, \CRANpkg{iptools}
-\citep{iptools} for handling IP addresses, and \CRANpkg{rgeolocate}
-\citep{rgeolocate} for direct IP geolocation.
+packages designed to overcome them: reading access logs in, manipulating
+URLs, manipulating IP addresses, and direct IP geolocation.
 
@@
 To make reading access logs into R as easy as possible we created the
-\pkg{webreadr} package. This contains user-friendly equivalents to
-\code{read.table} for each type of log, detecting the fields that should
-appear, converting the timestamps into POSIX objects, and merging fields
-or splitting fields where necessary. The package contains four core
-functions, one for each form of log: \code{read\_clf} for the Common Log
-Format, \code{read\_combined} for the Combined Log Format, and
-\code{read\_squid} and \code{read\_aws} for Squid and AWS formats
-respectively. Each one abstracts away the complexity of specifying
-column names and formats, and instead allows a researcher to read a file
-in with a minimal amount of work.
+\CRANpkg{webreadr} \citep{webreadr} package. This contains user-friendly
+equivalents to \code{read.table} for each type of log, detecting the
+fields that should appear, converting the timestamps into POSIX objects,
+and merging fields or splitting fields where necessary. The package
+contains four core functions, one for each form of log: \code{read\_clf}
+for the Common Log Format, \code{read\_combined} for the Combined Log
+Format, and \code{read\_squid} and \code{read\_aws} for Squid and AWS
+formats respectively. Each one abstracts away the complexity of
+specifying column names and formats, and instead allows a researcher to
+read a file in with a minimal amount of work.
 
@@
 
-To solve this common problem in analysing request logs, the
-\pkg{urltools} package was created. This contains a function,
-\code{url\_decode}, which decodes URLs - but relies on vectorised,
-compiled code. Benchmarking the two approaches against each other shows
-that the \pkg{urltools} implementation is approximately 60-70 times
-faster over large datasets. Again using \pkg{microbenchmark}, if we
-compare the vectorised decoding of 1,000,000 URLs with \pkg{urltools}
-against \code{URLdecode} in a \code{vapply} loop, we see a 60-70 times
-speed improvement:
+To solve this common problem in analysing request logs,
+the\CRANpkg{urltools} \citep{urltools} package was created. This
+contains a function, \code{url\_decode}, which decodes URLs and relies
+on vectorised, compiled code to do so. Benchmarking the two approaches
+against each other shows that the \pkg{urltools} implementation is
+approximately 60-70 times faster over large datasets. Again using
+\pkg{microbenchmark}, if we compare the vectorised decoding of 1,000,000
+URLs with \pkg{urltools} against \code{URLdecode} in a \code{vapply}
+loop, we see a 60-70 times speed improvement:
 
@@
 Based around the \emph{Boost ASIO} C++ library \citep{boostcite},
-\pkg{iptools} is a package designed for this kind of IP manipulation
-(and more). Built around combined code, it is extremely fast (for that
-code that does not rely on internet connections) and boasts a range of
-features.
+\CRANpkg{iptools} \citep{iptools} is a package designed for this kind of
+IP manipulation (and more). Built around combined code, it is extremely
+fast (for that code that does not rely on internet connections) and
+boasts a range of features.
 
@@
 
-With these concerns in mind, we wrote the \pkg{rgeolocate} package.
-Through \pkg{httr} this contains convenient, vectorised bindings to
-various web services that provide geographic metadata about IP
-addresses. More importantly, using the \CRANpkg{Rcpp} package to
-integrate C++ and R, \pkg{rgeolocate} also features a direct, compiled
-binding to the MaxMind API. This means that local binary databases can
-also be queried, which is far faster and more robust than web-based
-equivalents and avoids the privacy concerns associated with transmitting
-users' IP addresses externally.
+With these concerns in mind, we wrote the \CRANpkg{rgeolocate}
+\citep{rgeolocate} package. Through \pkg{httr} this contains convenient,
+vectorised bindings to various web services that provide geographic
+metadata about IP addresses. More importantly, using the \CRANpkg{Rcpp}
+package to integrate C++ and R, \pkg{rgeolocate} also features a direct,
+compiled binding to the MaxMind API. This means that local binary
+databases can also be queried, which is far faster and more robust than
+web-based equivalents and avoids the privacy concerns associated with
+transmitting users' IP addresses externally.
 
@@
 Wikimedia Foundation\\
-line 1\\ line 2\\
+149 New Montgomery Street, 6th floor\\ San Francisco, CA, 94105, USA\\
 }
-\href{mailto:author1@work}{\nolinkurl{author1@work}}
+\href{mailto:ironholds@gmail.com}{\nolinkurl{ironholds@gmail.com}}
 

### 2016-01-01 paper/paper.Rmd
update to include latest iptools changes

@@
 
-\pkg{iptools} also contains (for the same purpose) code to identify the actual client's IP address.
+Along with this general tool there are also functions for ascertaining specific facts about IP addresses, returning logical (\code{TRUE} or \code{FALSE}) values.
+These are \code{is\_valid}, \code{is\_ipv4}, \code{is\_ipv6}, which are built on top of \code{ip\_classify}, and \code{is\_multicast}, which identifies if an IP address is \dfn{multicast} - designed to point to multiple machines.
+
+\pkg{iptools} also contains (for similar purposes) code to identify the actual client's IP address.
 Access requests usually contain not only an IP address but a \dfn{X-Forwarded-For} field - a field identifying which \emph{other} IP addresses the request passed through, if the user who made the request is using some kind of proxy.
@@
 
-The limitation of the package is that it does not yet work on Windows - building ASIO on that platform is a uniquely frustrating experience - and that some operations do not yet support IPv6 (since they require the storage of numbers bigger than R can currently handle).
+The limitation of the package is that it some operations do not yet support IPv6 (since they require the storage of numbers bigger than R can currently handle).
 

### 2016-01-01 paper/paper.tex
update to include latest iptools changes

@@
 
-\abstract{%
+\abstract{
 Web access logs contain information on HTTP(S) requests and form a key
@@
 
-\pkg{iptools} also contains (for the same purpose) code to identify the
+Along with this general tool there are also functions for ascertaining
+specific facts about IP addresses, returning logical (\code{TRUE} or
+\code{FALSE}) values. These are \code{is\_valid}, \code{is\_ipv4},
+\code{is\_ipv6}, which are built on top of \code{ip\_classify}, and
+\code{is\_multicast}, which identifies if an IP address is
+\dfn{multicast} - designed to point to multiple machines.
+
+\pkg{iptools} also contains (for similar purposes) code to identify the
 actual client's IP address. Access requests usually contain not only an
@@
 
-The limitation of the package is that it does not yet work on Windows -
-building ASIO on that platform is a uniquely frustrating experience -
-and that some operations do not yet support IPv6 (since they require the
-storage of numbers bigger than R can currently handle).
+The limitation of the package is that it some operations do not yet
+support IPv6 (since they require the storage of numbers bigger than R
+can currently handle).
 
@@
 
-\address{%
+\address{
 Oliver Keyes\\
@@
 
-\address{%
+\address{
 Bob Rudis\\
@@
 
-\address{%
+\address{
 Jay Jacobs\\

### 2016-01-01 paper/paper.Rmd
made the images show up correctly

@@
   - name: Jay Jacobs
-    affiliation: Magical Python Land
+    affiliation: BitSight
     address:
-    - line 1
-    - line 2
-    email:  author3@work
+    - Boston, MA
+    email:  jay@beechplane.com
 abstract: >
@@
     \centering
-    \includegraphics[scale=0.4]{reading_benchmarks}
+    \includegraphics[width=14cm,keepaspectratio]{reading_benchmarks-jay}
     \caption{Results of microbenchmark run: read\char`_squid versus base-R equivalent code}
@@
     \centering
-    \includegraphics[scale=0.4]{decoding_benchmarks}
+    \includegraphics[width=14cm,keepaspectratio]{decoding_benchmarks-jay}
     \caption{Results of microbenchmark run: \code{url\_decode} versus base-R equivalent code}
@@
     \centering
-    \includegraphics[scale=0.4]{parsing_benchmarks}
+    \includegraphics[width=14cm,keepaspectratio]{parsing_benchmarks-jay}
     \caption{Results of microbenchmark run: \code{url\_parse} versus httr's equivalent code}

### 2016-01-01 paper/paper.tex
made the images show up correctly

@@
 \def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
+\tightlist
 \item
@@
 
-\address{
+\address{%
 Jay Jacobs\\
-Magical Python Land\\
+BitSight\\
 line 1\\ line 2\\
 }
-\href{mailto:author3@work}{\nolinkurl{author3@work}}
+\href{mailto:jjacobs@bitsighttech.com}{\nolinkurl{jjacobs@bitsighttech.com}}

### 2016-01-01 urltools_benchmarks.R

@@ -73,8 +73,9 @@ gg <- ggplot(dcode, aes(expr, time.sec)) +
         panel.grid.major=element_line(color="gray75"),
         panel.grid.minor=element_blank(),
         axis.title.y=element_blank(),
-        axis.ticks=element_blank())
-ggsave(gg, file = "./paper/decoding_benchmarks-jay.png")
+        axis.ticks=element_blank(),
+        title=element_text(size=10))
+ggsave(gg, file = "./paper/decoding_benchmarks-jay.png", width=7, height=3)
        
 
 ## parsing_benchmarks
@@ -94,6 +95,7 @@ gg <- ggplot(pcode, aes(expr, time.sec)) +
         panel.grid.major=element_line(color="gray75"),
         panel.grid.minor=element_blank(),
         axis.title.y=element_blank(),
-        axis.ticks=element_blank())
+        axis.ticks=element_blank(),
+        title=element_text(size=10))
 gg
-ggsave(gg, file = "./paper/parsing_benchmarks-jay.png")
+ggsave(gg, file = "./paper/parsing_benchmarks-jay.png", width=7, height=3)

### 2016-01-01 webreadr_benchmarks.R

@@ -53,6 +53,7 @@ gg <- ggplot(rdcode, aes(expr, time.sec)) +
         panel.grid.major=element_line(color="gray75"),
         #panel.grid.minor=element_blank(),
         axis.title.y=element_blank(),
-        axis.ticks=element_blank())
+        axis.ticks=element_blank(),
+        title=element_text(size=10))
 gg
-ggsave(gg, file = "./paper/reading_benchmarks-jay.png")
+ggsave(gg, file = "./paper/reading_benchmarks-jay.png", width=7, height=3)

### 2016-01-02 paper/paper.Rmd
Update paper.Rmd
typo: missing space to_include

@@
 ### Decoding
-Both values can be percent-encoded, allowing them toinclude characters that are valid but reserved by the URL specification as having special meanings ("reserved characters").
+Both values can be percent-encoded, allowing them to include characters that are valid but reserved by the URL specification as having special meanings ("reserved characters").
 A \samp{\#} symbol, for example, is encoded as \samp{\%23}: a percentage symbol, followed by the byte-code for that character.

### 2016-01-02 paper/paper.Rmd
Update paper.Rmd
Added missing space
Update paper.Rmd
typo

@@
 
-To solve this common problem in analysing request logs, the\CRANpkg{urltools} \citep{urltools} package was created. This contains a function,
+To solve this common problem in analysing request logs, the \CRANpkg{urltools} \citep{urltools} package was created. This contains a function,
 \code{url\_decode}, which decodes URLs and relies on vectorised, compiled code to do so.
@@
 
-The limitation of the package is that it some operations do not yet support IPv6 (since they require the storage of numbers bigger than R can currently handle).
+The limitation of the package is that some operations do not yet support IPv6 (since they require the storage of numbers bigger than R can currently handle).
 
